---
title: "MDCEV Exploration"
author: "Zander Gordan"
date: "2/27/2021"
output: html_document
runtime: shiny
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document provides some exploration of the Multiple-Discrete Continuous
Extreme Value (MDCEV) model. It follows the notation in [this R Journal article](https://journal.r-project.org/archive/2021/RJ-2021-015/index.html) 
for the package `rmdcev`, authored by
[Patrick Lloyd-Smith](https://plloydsmith.github.io). 

The code for this document is on Github [here](https://github.com/zandergordan/shiny-server/blob/main/mdcev_notebook/mdcev_notebook.rmd).

## The MDCEV Utility function

The MDCEV model is focused on predicting consumption of a fixed set of goods, 
such as trips to different national parks. The general form of the utility 
function for this model is:

$$ U(x_1, x_2, \ldots, x_K) = \sum_{k=2}^{K} \frac{\gamma_k}{\alpha_k} \psi_k 
\left[ \left( \frac{x_k}{\gamma_k} + 1 \right)^{\alpha_k} - 1 \right] + 
\frac{\psi_1}{\alpha_1} x_1^{\alpha_1} $$

Where $x_2$ through $x_k$ represent the quantity consumed of the goods we are
trying to model, and $x_1$ represents the quantity consumed of the numeraire
good. 

In order for the model to behave, we restrict $\gamma_k > 0$, $\psi_k > 0$ and 
$\alpha_k \le 1$, $\forall k$.

There are 4 different special cases of this model which can be estimated by the 
`rmdcev` package:

1. $\alpha$-profile, with $\gamma_k=1$, $\forall k$.

$$ U(x_1, x_2, \ldots, x_K) = \sum_{k=2}^{K} \frac{\psi_k}{\alpha_k} 
\left[ \left( x_k + 1 \right)^{\alpha_k} - 1 \right] + 
\frac{\psi_1}{\alpha_1} x_1^{\alpha_1} $$

2. $\gamma$-profile, with $\alpha_k=0$, $\forall k > 1$.

$$ U(x_1, x_2, \ldots, x_K) = \sum_{k=2}^{K} \gamma_k \psi_k 
 \ln \left(\frac{x_k}{\gamma_k} + 1 \right) + 
\frac{\psi_1}{\alpha_1} x_1^{\alpha_1} $$

```{r}
ux <- function(x, psi, gamma, alpha) {
        psi * ((x + gamma) / gamma)^(alpha - 1)
    }
```



```{r MDCEV_viz, echo=FALSE}
inputPanel(
  sliderInput("psi", label = "Value of Psi parameter",
              min = 1, max = 50, value = 30),
  
  sliderInput("gamma", label = "Value of Gamma",
              min = 0.1, max = 10, value = 2),
  
  sliderInput("alpha", label = "Value of Alpha parameter",
              min = 0, max = 1, value = 0.5)
)

renderPlot({
  # draw the marginal utility curve with specified parameters
  plot(seq(0, 10, 0.1),
       ux(seq(0, 10, 0.1), psi = input$psi, gamma = input$gamma, alpha = input$alpha),
       type = "l")
    })
```
